RationalGRL: A Framework for Argumentation and Goal Modeling

### 2017-07-10 overview.tex
Section 3.1 and 3.2 has been rewritten

In this section we present a high-level overview of the RationalGRL framework. In the first subsection we present a methodology, specifying how practitioners can use RationalGRL to create goal models with traceability links to underlying arguments. In the second subsection we link the new argumentation elements and relations to the existing elements and relations of GRL in a metamodel. This metamodel serves as the specification for an implementation. In Section~\ref{sect:tool} we briefly discuss our prototype implementation based on this metamodel.

@@

To develop the argumentation model, we need to analyze critical questions, argument schemes and the stakeholder's discussions. For the GRL part of the framework, we first need to create the "initial" GRL models by analyzing the non-functional requirements in the requirements specification document and by refining the high-level goals into operationalized tasks. These two parts, GRL and argumentation models are developed iteratively and each side can impact the other side so that the models can be refined or new critical questions and argument schemes can be instantiated. For example, answering a critical question \emph{Is the task \texttt{A} possible?}, instantiated from the argumentation model, can result in removing or adding a task in the GRL model. Similarly,  if, for example, we add a new intentional element to the GRL model, it can lead to a new critical question relevant to this intentional element and its relationships with other intentional elements. Figure~\ref{fig:rationalgrl-framework} presents an overview of RationalGRL framework with its components and their relationships.  The GRL models  are shown  on the right-hand side of the framework while the argumentation models is on the left-hand side. The links between the two sides illustrates the impacts and relationships between two sides.

@@

We propose the following methodology (shown in Figure~\ref{fig:rationalgrl-methodology}) to develop an instance of the RationalGRL framework. Note that, here we assume that the initial GRL models have been created based on the requirements specification documents and the discussions of the stakeholders. The rest of the steps are as follows:

@@

%TODO I need to reword up to the example.

\textbf{Instantiate Argument Schemes (AS)} -- After creating the initial GRL models, we In the first step, we instantiate one of the AS from the argumentation framework. In the section we develop 13 arguments schemes which we use in our analysis. An example of an argument scheme is "Goal \emph{G} contributes to softgoal \emph{S}". An instantiation of an argument scheme thus corresponds to an argument for or against part of a goal model.

\textbf{Answer Critical Questions (CQs)} -- Instantiated arguments can be attacked by counter-arguments. CQs are ways in which AS can be attacked. Each argument scheme includes one or more CQs. For example, for the argument scheme mentioned above, we have two CQs as: \emph{Does the goal contributes to the softgoal?} and \emph{Does the goal contributes to some other softgoals?}. Note that, the answer to CQ can also result into ``conflict'' situation which we do not consider here. Answering a CQ may result in an instantiation of a new AS. Thus, it is possible to go back and forth between this step and the previous one.

\textbf{Decide on IE and the Relationships} -- The answers to the CQs can result in one of the four cases which impact the arguments and corresponding GRL IE: \textsf{INTRO}, \textsf{DISABLE}, \textsf{REPLACE} or \textsf{ATTACK}.  \textsf{INTRO} means that the argument scheme of the CQ does not get attacked and instead it creates a new argument. \textsf{DISABLE} means that the IE or the relationship related to the AS needs to be disabled or removed from the models. \textsf{REPLACE} introduces a new argument and attack the original argument at the same time. It basically replace the original element of the AS with a new one.  \textsf{ATTACK} is a generic counterargument which attacks any argument with another argument when a new evidence occur.

\textbf{Modify GRL Models} -- Based on the result of step three, the GRL models can be modified. In this case, either a new IE or a new relationships is introduced, an existing IE or relationship gets disabled (removed) from the model or finally an existing IE or relationships is replaced by a new one. This results in a new modified GRL. The new GRL model can then impact the argument schemes and instantiate another argument scheme.

@@

\subsubsection*{Example} \todo{Sepideh}{Marc}{Do we need an example here? I think no...this is from istar also, if we want it, we need to modify it}
\label{sect:methodology-example}

Our examples comes from a transcript containing discussions about the development of an information system, and they are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}. We provide several more examples in Section~\ref{sect:examples}.

A group of stakeholders is developing a goal model for a traffic simulator example and they are modeling the goal \emph{simulate} of the system using the RationalGRL methodology. This proceeds in the following way:
\begin{itemize}
\item
First they start at step \emph{Modify GRL models} (Figure~\ref{fig:rationalgrl-methodology}), and add the IE \texttt{Simulate} to GRL model (Figure~\ref{fig:examples:decomposition}, GRL model, top IE).
\item Next, they switch to the argumentation part (step \emph{Instantiate arguments schemes}). They answer CQ \emph{Does Simulate AND-decompose into any tasks?} with \emph{Yes: namely tasks ``Dynamic simulation'' and ``Static simulation}.
\item As a result, two argument schemes are created, namely:
\begin{itemize}
\item Actor \emph{System} has task \emph{Dynamic simulation}
\item Actor \emph{System} has task \emph{Static simulation}
\end{itemize}
\item In the GRL model, this corresponds to the addition of two tasks, and an AND-decomposition from the goal \emph{Simulate} to these two tasks.
\item Next, the stakeholders test the validity of their goal model by answering more CQ. They answer two CQs:
\begin{itemize}
\item
CQ \emph{Is task ``Dynamic simulation'' relevant} is answered with ``No, it is not relevant since the problem specification explicitly states dynamic simulations are not required''. Thus, the corresponding GRL IE is disabled.
\item The decomposition is changed from AND to OR, since it turned out not both tasks can be implemented together.
\end{itemize}
\end{itemize}

>>>>

In this section we present a high-level overview of the RationalGRL framework. In the first subsection, we present a methodology, specifying how practitioners can use RationalGRL to create goal models with traceability links to underlying arguments. In the second subsection, we link the new argumentation elements and relations to the existing elements and relations of GRL in a metamodel. This metamodel serves as the specification for an implementation. In Section~\ref{sect:tool} we briefly discuss our prototype implementation based on this metamodel.

@@

To develop the argumentation model, we need to analyze critical questions, argument schemes and the stakeholder's discussions. For the GRL part, we first need to create the "initial" GRL models by analyzing the non-functional requirements in the requirements specification document and by refining the high-level goals into operationalized tasks. These two parts, GRL and argumentation models are developed iteratively and each side can impact the other side so that the models can be refined or new critical questions and argument schemes can be instantiated. For example, answering a critical question \emph{Is the task \texttt{A} possible?}, instantiated from the argumentation model, can result in removing or adding a task in the GRL model. Similarly,  if, for example, we add a new intentional element to the GRL model, it can lead to a new critical question relevant to this intentional element and its relationships.  Figure~\ref{fig:rationalgrl-framework} presents an overview of RationalGRL framework with its components and their relationships.  The GRL models  are shown  on the right-hand side of the framework while the argumentation models is on the left-hand side. The links between the two sides illustrates the impacts and relationships between two sides. Note that, any of the critical question, their answer and the argument schemes that are instantiated during the analysis phase of the GRL model are documented with the GRL model and can be referred to in the future.

@@

We propose the following methodology (shown in Figure~\ref{fig:rationalgrl-methodology}) to develop an instance of the RationalGRL framework. Here we assume that the initial GRL models have been created based on the requirements specification documents and the discussions of the stakeholders. The rest of the steps are as follows:

@@

\textbf{(1) Instantiate Argument Schemes (AS)} -- In this step, we start from the list of arguments schemes of the argumentation framework. We select an intentional element from the initial GRL model that we want to analyze and we instantiate a relevant argument scheme from the already existing list of argument schemes or by adding a new one. For example, an argument scheme can be "Goal \emph{G} contributes to softgoal \emph{S}". When an argument scheme is instantiated, it corresponds to  an argument for or against part of a goal model.

\textbf{(2) Answer Critical Questions (CQs)} -- After instantiating an argument scheme, we invoke related critical questions o attack the argument by counter-arguments.  Each argument scheme includes one or more critical question. For example, for the argument scheme, "Goal \emph{G} contributes to softgoal \emph{S}", there are two critical questions as follows:  \emph{Does the goal contributes to the softgoal?} and \emph{Does the goal contributes to some other softgoals?}. It is worth mentioning that, answering a critical question may result in a "conflict" situation. This is out of the scope of our work here.  When the analyst answers  a critical question, a new argument scheme may be instantiated.  Thus, it is possible to go back and forth between this step and the step (1).

\textbf{(3) Decide on Intentional Elements and their Relationships} -- By answering a critical question, one of the four following cases can occur: \textsf{INTRO}, \textsf{DISABLE}, \textsf{REPLACE} or \textsf{ATTACK}.  Any of these cases can  impact the arguments and corresponding GRL intentional elements.  \textsf{INTRO} means that 
a new argument scheme is created. That means, the current argument scheme related to the critical question does not get attacked.  In the case of \textsf{DISABLE}, the intentional element or its related links are disabled or removed from the models. \textsf{REPLACE} introduces a new argument and attacks the original argument at the same time. This means that the original element of the argument scheme is replaced with a new one.   \textsf{ATTACK} is a generic counterargument which attacks any argument with another argument when a new evidence occur.

\textbf{(4) Modify GRL Models} -- In this step, we modify the GRL models based on the situation of step (3). That is, one of the following situation can happen with respect to the initial GRL model: 1) a new intentional element or a new link is introduced; 2) an existing intentional element or an existing link gets disabled (removed) from the model; or 3) an existing intentional element or link is replaced by a new one. This results in a new modified GRL. The new GRL model can then impact the argument schemes and instantiate another argument scheme (Step (1)).

We can continue these four steps until there is no more intentional element or link to analyze or we reach a satisfactory model.

@@

\subsubsection*{Example} \todo{Sepideh}{Marc}{Is this part ok? I am not sure if we should have this section at all here. It may make things more complicated. Better to have it all in the example section}
\label{sect:methodology-example}

In this section, we present a small example for the RationalGRL methodology. This example, which is based on the traffic simulator example described in Section~\ref{sect:goals:runningexample}. We base the argument schemes and the critical questions of this example on a transcript containing discussions about the development of this traffic simulator system. These transcripts are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}. We provide several more examples in Section~\ref{sect:examples}.

As mentioned in Section~\ref{sect:goals:runningexample}, a group of stakeholders is developing a goal model for a traffic simulator example and they are modeling the goal \texttt{Simulate Traffic} using the RationalGRL methodology. This proceeds in the following way:
\begin{itemize}
\item
First they start %at step \emph{Modify GRL models} (Figure~\ref{fig:rationalgrl-methodology}), 
with the initial GRL model (Figure~\ref{fig:example-small})  and they invoke the intentional element \texttt{Simulate Traffic}. %SG: We say above that we have an initial GRL model, so how do we match it with this sentence? I tried to reword it but it doesn't make so much sense... 
\item Next, they switch to the argumentation part (step \emph{Instantiate arguments schemes}). They answer critical question \emph{Does Simulate Traffic AND-decompose into any tasks?} with \emph{Yes: namely tasks ``Dynamic Simulation'' and ``Static Simulation}.
\item As a result, two argument schemes are created, namely:
\begin{itemize}
\item Actor \texttt{Traffic Simulator} has task \texttt{Dynamic Simulation}
\item Actor \texttt{Traffic Simulator} has task \texttt{Static Simulation}
\end{itemize}
\item In the GRL model, this corresponds to the addition of two tasks, and an AND-decomposition from the goal \emph{Simulate Traffic} to these two tasks.
\item Next, the stakeholders test the validity of their goal model by answering more critical question. They answer two critical questions:
\begin{itemize}
\item
CQ \emph{Is task ``Dynamic Simulation'' relevant} is answered with ``No, it is not relevant since the problem specification explicitly states dynamic simulations are not required''. Thus, the corresponding GRL intentional element is disabled.
\item The decomposition is changed from AND to OR, since it turned out not both tasks can be implemented together.
\end{itemize}
\end{itemize}
